---
title: "NJ_CA_comparison"
author: "Claire Morton"
date: "11/10/2022"
output: html_document
---

This script will compare the thresholds in the New Jersey law with the potential thresholds for the California law, to try to 
understand how comparable the two are (and how to set up potential CA thresholds that are more comparable to CES)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Load data

```{r load_data}
nj <- read_csv("../../data/data_interim/NJ_census_block_group.csv")
ca <- read_csv("../../data/data_interim/CA_census_block_group.csv")
```

```{r thresh}
ca_pov <- .35
nj_pov <- .35

ca_lep <- .4
nj_lep <- .4

ca_raceeth <- .75
nj_raceeth <- .4
```

```{r indicators}
ca <- ca %>%
  mutate(raceeth_indicator = ifelse((!is.na(propHispanicOrNonWhite) & propHispanicOrNonWhite >= ca_raceeth), 1, 0),
         pov_indicator = ifelse((!is.na(propPoverty) & propPoverty >= ca_pov), 1, 0),
         lep_indicator = ifelse((!is.na(propLEP) & propLEP >= ca_lep), 1, 0))

nj <- nj %>%
  mutate(raceeth_indicator = ifelse((!is.na(propHispanicOrNonWhite) & propHispanicOrNonWhite >= nj_raceeth), 1, 0),
         pov_indicator = ifelse((!is.na(propPoverty) & propPoverty >= nj_pov), 1, 0),
         lep_indicator = ifelse((!is.na(propLEP) & propLEP >= nj_lep), 1, 0))
```

What % of block groups are classified as DACs, and how does that break down into % of block groups classified as each of the three things?
```{r analysis_1}
# % of block groups classified as DACs
paste0(round(sum(ca$univariate_DAC)/nrow(ca)*100, 2), "% of block groups in California (", sum(ca$univariate_DAC), "), are classified as DACs")
paste0(round(sum(nj$univariate_DAC)/nrow(nj)*100, 2), "% of block groups in New Jersey (", sum(nj$univariate_DAC), "), are classified as DACs")

# % of block groups classified as DACs based on poverty
paste0(round(sum(ca$pov_indicator)/nrow(ca)*100, 2), "% of block groups in California (", sum(ca$pov_indicator), "), are above the poverty threshold")
paste0(round(sum(nj$pov_indicator)/nrow(nj)*100, 2), "% of block groups in New Jersey (", sum(nj$pov_indicator), "), are above the poverty threshold")

# % of block groups classified as DACs based on race/ethnicity
paste0(round(sum(ca$raceeth_indicator)/nrow(ca)*100, 2), "% of block groups in California (", sum(ca$raceeth_indicator), "), are above the race/ethnicity threshold")
paste0(round(sum(nj$raceeth_indicator)/nrow(nj)*100, 2), "% of block groups in New Jersey (", sum(nj$raceeth_indicator), "), are above the race/ethnicity threshold")

# % of block groups classified as DACs based on race/ethnicity
paste0(round(sum(ca$lep_indicator)/nrow(ca)*100, 2), "% of block groups in California (", sum(ca$lep_indicator), "), are above the linguistic isolation threshold")
paste0(round(sum(nj$lep_indicator)/nrow(nj)*100, 2), "% of block groups in New Jersey (", sum(nj$lep_indicator), "), are above the linguistic isolation threshold")
```

What would thresholds in the same quantile be for CA? 
```{r analysis_2}
ca_pov_quant <- quantile(ca$propPoverty, 1-c(sum(nj$pov_indicator)/nrow(nj)), na.rm = T)
paste0("If we set a threshold for poverty in California to capture the same percentage of block",
       " groups as New Jersey's threshold, we would set the threshold at ", round(ca_pov_quant, 2),
      " instead of ", ca_pov)

ca_raceeth_quant <- quantile(ca$propHispanicOrNonWhite, 1-c(sum(nj$raceeth_indicator)/nrow(nj)), na.rm = T)
paste0("If we set a threshold for race/ethnicity in California to capture the same percentage of block",
       " groups as New Jersey's threshold, we would set the threshold at ", round(ca_raceeth_quant, 2),
       " instead of ", ca_raceeth)

ca_lep_quant <- quantile(ca$propLEP, 1-c(sum(nj$lep_indicator)/nrow(nj)), na.rm = T)
paste0("If we set a threshold for linguistic isolation in California to capture the same percentage of block",
       " groups as New Jersey's threshold, we would set the threshold at ", round(ca_lep_quant, 2),
       " instead of ", ca_lep)
```

How far off the mean are thresholds
```{r analysis_3}
# mean poverty, and difference to mean poverty in DACs
paste0("In California, mean poverty level is ", round(mean(ca$propPoverty, na.rm = T), 2), ", and mean poverty in DACs is ",
       round(mean((ca%>%filter(univariate_DAC==1))$propPoverty, na.rm = T), 2))
paste0("In New Jersey, mean poverty level is ", round(mean(nj$propPoverty, na.rm = T), 2), ", and mean poverty in DACs is ",
       round(mean((nj%>%filter(univariate_DAC==1))$propPoverty, na.rm = T), 2))

# mean raceeth, and difference to mean raceeth in DACs
paste0("In California, mean proportion of Hispanic or non-white people is ", round(mean(ca$propHispanicOrNonWhite, na.rm = T), 2), ", and mean proportion of Hispanic or non-white people in DACs is ",
       round(mean((ca%>%filter(univariate_DAC==1))$propHispanicOrNonWhite, na.rm = T), 2))
paste0("In New Jersey, mean proportion of Hispanic or non-white people is ", round(mean(nj$propHispanicOrNonWhite, na.rm = T), 2), ", and mean proportion of Hispanic or non-white people in DACs is ",
       round(mean((nj%>%filter(univariate_DAC==1))$propHispanicOrNonWhite, na.rm = T), 2))

# mean LEP, and difference to mean LEP in DACs
paste0("In California, mean linguistic isolation is ", round(mean(ca$propLEP, na.rm = T), 2), ", and mean linguistic isolation in DACs is ",
       round(mean((ca%>%filter(univariate_DAC==1))$propLEP, na.rm = T), 2))
paste0("In New Jersey, mean linguistic isolation is ", round(mean(nj$propLEP, na.rm = T), 2), ", and mean linguistic isolation in DACs is ",
       round(mean((nj%>%filter(univariate_DAC==1))$propLEP, na.rm = T), 2))

```

How far off the median are thresholds
```{r analysis_3}
# median poverty, and difference to median poverty in DACs
paste0("In California, median poverty level is ", round(median(ca$propPoverty, na.rm = T), 2), ", and median poverty in DACs is ",
       round(median((ca%>%filter(univariate_DAC==1))$propPoverty, na.rm = T), 2))
paste0("In New Jersey, median poverty level is ", round(median(nj$propPoverty, na.rm = T), 2), ", and median poverty in DACs is ",
       round(median((nj%>%filter(univariate_DAC==1))$propPoverty, na.rm = T), 2))

# median raceeth, and difference to median raceeth in DACs
paste0("In California, median proportion of Hispanic or non-white people is ", round(median(ca$propHispanicOrNonWhite, na.rm = T), 2), ", and median proportion of Hispanic or non-white people in DACs is ",
       round(median((ca%>%filter(univariate_DAC==1))$propHispanicOrNonWhite, na.rm = T), 2))
paste0("In New Jersey, median proportion of Hispanic or non-white people is ", round(median(nj$propHispanicOrNonWhite, na.rm = T), 2), ", and median proportion of Hispanic or non-white people in DACs is ",
       round(median((nj%>%filter(univariate_DAC==1))$propHispanicOrNonWhite, na.rm = T), 2))

# median LEP, and difference to median LEP in DACs
paste0("In California, median linguistic isolation is ", round(median(ca$propLEP, na.rm = T), 2), ", and median linguistic isolation in DACs is ",
       round(median((ca%>%filter(univariate_DAC==1))$propLEP, na.rm = T), 2))
paste0("In New Jersey, median linguistic isolation is ", round(median(nj$propLEP, na.rm = T), 2), ", and median linguistic isolation in DACs is ",
       round(median((nj%>%filter(univariate_DAC==1))$propLEP, na.rm = T), 2))

```

For discussion:
- How to make this more comparable to CES given that NJ has a very similar proportion of block groups being classified as DACs?

